---
title: "Statistical analysis of cerebellar tremor frequency"
author: 'Sofia ZEMMAHI '
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output:
  html_document:
    df_print: paged
    code_folding: hide
    toc: yes
    number_sections: yes
  pdf_document:
    toc: yes
---

# Objective

To verify if a smartphone accelerometre *PhyPhox* can measure cerebellar tremor reproducibly.

And More ore specifically:
Are the obtained frequencies consistent:

* between repetitions,
* between days,
* and between conditions?

The main variable of this study is the dominant frequency because cerebellar tremor is defined by a low frequency (3-5 Hz),
and the use of FFT to extract that frequency.

Here the main outcome is Dominant_Frequency_Hz.

A secondary essential variable is RMS wich is the Root Mean square amplitude. 
It measures the intensity of the tremor, not the frequent.
It is used to: **quantify the intensity, verify that the tremor increases during the action and strengthen clinical consistency.**

So the main question of this analysis is : 
**Does the smartphone *PhyPhox* accelerometre measure a coherent frequency with cerebellar tremor?**

We can answer this principal question in three under-parts : 
1) Do the frequencies remain stable between repetitions? -> Intra day reproducibility.
2) Do the frequencies remain consistent between days? -> Inter day reproducibility.
3) Does tremor increase during action?


# Part 1 : Intra-day reproducibility

## Do dominant tremor frequencies remain stable between repetitions?

The objective of this first part is to evaluate short-term reproducibility of dominant tremor frequency measurements obtained using smartphone accelerometry and FFT analysis.
For each recording day, dominant tremor frequency was analyzed across repeated measurements performed during rest and task conditions.

A stable frequency between repetitions would support the reproducibility of the measurement method.

# 1.1 Day 1 analysis

## 1.1.1 Descriptive statistics

Objective: To describe dominant tremor frequencies obtained during **Day 1**.

We calculated:
- mean frequency,
- standard deviation (SD),
- median,
- minimum frequency,
- maximum frequency.

These descriptive statistics help verify whether the obtained frequencies remain within the classical cerebellar tremor range reported in the literature (3–5 Hz).

```{r}
# Select Day 1 data
day1_data <- all_results %>%
  filter(Day == "D1")
# Descriptive statistics for Day 1
day1_stats <- day1_data %>%
  group_by(Condition) %>%
  summarise(
    mean_frequency = mean(Dominant_Frequency_Hz),
    sd_frequency = sd(Dominant_Frequency_Hz),
    median_frequency = median(Dominant_Frequency_Hz),
    min_frequency = min(Dominant_Frequency_Hz),
    max_frequency = max(Dominant_Frequency_Hz),
    .groups = "drop"
  )
day1_stats
```

Dominant tremor frequency during Day 1 remained within the classical cerebellar tremor range (3–5 Hz) for both rest and task conditions.



1.1.2 Dominant frequency evolution across repetitions

Objective: To evaluate frequency stability across repetitions during Day 1.

A line plot was used to visualize frequency evolution over repeated measurements.

```{r}
ggplot(day1_data,
       aes(x = Repetition,
           y = Dominant_Frequency_Hz,
           color = Condition,
           group = Condition)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 4) +
  geom_text(
    aes(label = round(Dominant_Frequency_Hz, 2)),
    vjust = -1,
    size = 4,
    color = "black"
  ) +
  scale_color_manual(values = c(
    "Rest" = "lightblue",
    "Task" = "salmon"
  )) +
  scale_x_continuous(breaks = c(1,2,3,4)) +
  coord_cartesian(ylim = c(3,5)) +
  labs(
    title = "Day 1 - Dominant tremor frequency across repetitions",
    x = "Repetition",
    y = "Dominant frequency (Hz)",
    color = "Condition"
  ) +
  theme_minimal(base_size = 10)
```


During task conditions, dominant tremor frequency remained relatively stable across repetitions (3.72 Hz then 3.63 Hz), suggesting good short-term reproducibility.

Greater variability was observed during rest conditions, with frequencies increasing from 3.27 Hz to 4.69 Hz and 4.85 Hz across repetitions.

This variability may reflect transient postural oscillations, movement artefacts, or residual oscillatory activity after movement.




1.1.3 RMS amplitude evolution across repetitions

Objective: To evaluate tremor intensity modulation during Day 1.

RMS amplitude reflects tremor intensity rather than tremor frequency.

```{r}
ggplot(day1_data,
       aes(x = Repetition,
           y = RMS_Amplitude,
           color = Condition,
           group = Condition)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 4) +
  geom_text(
    aes(label = round(RMS_Amplitude, 2)),
    vjust = -1,
    size = 4,
    color = "black"
  ) +
  scale_color_manual(values = c(
    "Rest" = "lightblue",
    "Task" = "salmon"
  )) +
  scale_x_continuous(breaks = c(1,2,3,4)) +
  coord_cartesian(ylim = c(0,1)) +
  labs(
    title = "Day 1 - RMS amplitude across repetitions",
    x = "Repetition",
    y = "RMS amplitude",
    color = "Condition"
  ) +
  theme_minimal(base_size = 10)

```

RMS amplitudes were substantially higher during task conditions (0.69–0.72) compared with rest periods (0.07–0.26), indicating increased tremor intensity during voluntary movement.

A transient RMS increase was observed during the second rest period, possibly reflecting residual post-movement oscillations or brief movement artefacts.



1.1.4 Coefficient of variation (CV)

Objective: To evaluate measurement variability across repetitions.

The coefficient of variation (CV) was calculated for:

* dominant tremor frequency,
* RMS amplitude.

A low CV indicates good reproducibility and low variability between repetitions.

```{r}
day1_cv <- day1_data %>%
  group_by(Condition) %>%
  summarise(
    mean_frequency = mean(Dominant_Frequency_Hz),
    sd_frequency = sd(Dominant_Frequency_Hz),
    CV_frequency = (sd_frequency / mean_frequency) * 100,
    mean_RMS = mean(RMS_Amplitude),
    sd_RMS = sd(RMS_Amplitude),
    CV_RMS = (sd_RMS / mean_RMS) * 100,
    .groups = "drop"
  )
day1_cv

```

During task conditions, dominant frequency showed a low coefficient of variation (CV = 1.73%), suggesting good reproducibility across repetitions.

Conversely, RMS amplitude during rest showed higher variability (CV = 82.27%), reflecting fluctuations in tremor intensity during resting periods.

1.1.5 Day 1 synthesis

Overall, Day 1 analysis suggests that smartphone accelerometry combined with FFT analysis provides coherent dominant tremor frequency measurements consistent with cerebellar tremor characteristics.

Dominant frequency remained within the classical cerebellar tremor range (3–5 Hz) and showed relatively good stability across task repetitions.

RMS analysis demonstrated that tremor intensity increased substantially during voluntary movement compared with rest conditions.

These findings support the feasibility of smartphone-based tremor frequency quantification during repeated measurements.





# 1.2 Day 2 analysis

## 1.2.1 Descriptive statistics

Objective: To describe dominant tremor frequencies obtained during Day 2.

We calculated:
- mean frequency,
- standard deviation (SD),
- median,
- minimum frequency,
- maximum frequency.

These descriptive statistics help verify whether the obtained frequencies remain within the classical cerebellar tremor range reported in the literature (3–5 Hz).

```{r}
# Select Day 2 data

day2_data <- all_results %>%
  filter(Day == "D2")

# Descriptive statistics for Day 2

day2_stats <- day2_data %>%
  group_by(Condition) %>%
  summarise(
    mean_frequency = mean(Dominant_Frequency_Hz),
    sd_frequency = sd(Dominant_Frequency_Hz),
    median_frequency = median(Dominant_Frequency_Hz),
    min_frequency = min(Dominant_Frequency_Hz),
    max_frequency = max(Dominant_Frequency_Hz),
    .groups = "drop"
  )

day2_stats
```
Dominant tremor frequency during Day 2 remained within the classical cerebellar tremor range (3–5 Hz) for both rest and task conditions.

1.2.2 Dominant frequency evolution across repetitions

Objective: To evaluate frequency stability across repetitions during Day 2.

A line plot was used to visualize frequency evolution over repeated measurements.
```{r}
ggplot(day2_data,

       aes(x = Repetition,

           y = Dominant_Frequency_Hz,

           color = Condition,

           group = Condition)) +

  geom_line(linewidth = 1.2) +

  geom_point(size = 4) +

  geom_text(

    aes(label = round(Dominant_Frequency_Hz, 2)),

    vjust = -1,

    size = 4,

    color = "black"

  ) +

  scale_color_manual(values = c(

    "Rest" = "lightblue",

    "Task" = "salmon"

  )) +

  scale_x_continuous(breaks = c(1,2,3,4)) +

  coord_cartesian(ylim = c(3,5)) +

  labs(

    title = "Day 2 - Dominant tremor frequency across repetitions",

    x = "Repetition",

    y = "Dominant frequency (Hz)",

    color = "Condition"

  ) +

  theme_minimal(base_size = 10)
```
Dominant tremor frequency during Day 2 remained globally stable across repetitions and conditions.

Task frequencies remained within the expected cerebellar tremor frequency range, supporting short-term reproducibility of frequency measurements.



1.2.3 RMS amplitude evolution across repetitions

Objective: To evaluate tremor intensity modulation during Day 2.

RMS amplitude reflects tremor intensity rather than tremor frequency.
```{r}
ggplot(day2_data,

       aes(x = Repetition,

           y = RMS_Amplitude,

           color = Condition,

           group = Condition)) +

  geom_line(linewidth = 1.2) +

  geom_point(size = 4) +

  geom_text(

    aes(label = round(RMS_Amplitude, 2)),

    vjust = -1,

    size = 4,

    color = "black"

  ) +

  scale_color_manual(values = c(

    "Rest" = "lightblue",

    "Task" = "salmon"

  )) +

  scale_x_continuous(breaks = c(1,2,3,4)) +

  coord_cartesian(ylim = c(0,1.2)) +

  labs(

    title = "Day 2 - RMS amplitude across repetitions",

    x = "Repetition",

    y = "RMS amplitude",

    color = "Condition"

  ) +

  theme_minimal(base_size = 10)
```
RMS amplitudes remained substantially higher during task conditions than during rest periods, indicating increased tremor intensity during voluntary movement.
This observation is consistent with the action-related characteristics of cerebellar tremor.


1.2.4 Coefficient of variation (CV)

Objective: To evaluate measurement variability across repetitions.

The coefficient of variation (CV) was calculated for:

* dominant tremor frequency,
* RMS amplitude.

A low CV indicates good reproducibility and low variability between repetitions.
```{r}
day2_cv <- day2_data %>%

  group_by(Condition) %>%

  summarise(

    mean_frequency = mean(Dominant_Frequency_Hz),

    sd_frequency = sd(Dominant_Frequency_Hz),

    CV_frequency = (sd_frequency / mean_frequency) * 100,

    mean_RMS = mean(RMS_Amplitude),

    sd_RMS = sd(RMS_Amplitude),

    CV_RMS = (sd_RMS / mean_RMS) * 100,

    .groups = "drop"

  )

day2_cv
```

The coefficient of variation remained relatively low for dominant frequency during task conditions, suggesting good reproducibility across repeated measurements.

Higher RMS variability was observed during rest periods compared with task conditions.


1.2.5 Day 2 synthesis

Overall, Day 2 analysis supports the reproducibility of smartphone-based accelerometry measurements for cerebellar tremor frequency quantification.

Dominant tremor frequency remained within the expected cerebellar tremor range and showed relatively stable values across repetitions.

RMS amplitude remained substantially higher during voluntary movement, supporting the clinical coherence of the measurements obtained using smartphone accelerometry and FFT analysis.



# 1.3 Day 3 analysis

## 1.3.1 Descriptive statistics

Objective: To describe dominant tremor frequencies obtained during Day 3.

We calculated:
- mean frequency,
- standard deviation (SD),
- median,
- minimum frequency,
- maximum frequency.

These descriptive statistics help verify whether the obtained frequencies remain within the classical cerebellar tremor range reported in the literature (3–5 Hz).

```{r}
# Select Day 3 data

day3_data <- all_results %>%
  filter(Day == "D3")

# Descriptive statistics for Day 3
day3_stats <- day3_data %>%
  group_by(Condition) %>%
  summarise(
    mean_frequency = mean(Dominant_Frequency_Hz),
    sd_frequency = sd(Dominant_Frequency_Hz),
    median_frequency = median(Dominant_Frequency_Hz),
    min_frequency = min(Dominant_Frequency_Hz),
    max_frequency = max(Dominant_Frequency_Hz),
    .groups = "drop"
  )

day3_stats
```

Dominant tremor frequency during Day 3 remained within the classical cerebellar tremor range (3–5 Hz) for both rest and task conditions.

1.3.2 Dominant frequency evolution across repetitions

Objective: To evaluate frequency stability across repetitions during Day 3.

A line plot was used to visualize frequency evolution over repeated measurements.

```{r}
ggplot(day3_data,
       aes(x = Repetition,
           y = Dominant_Frequency_Hz,
           color = Condition,
           group = Condition)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 4) +
  geom_text(
    aes(label = round(Dominant_Frequency_Hz, 2)),
    vjust = -1,
    size = 4,
    color = "black"
  ) +
  scale_color_manual(values = c(
    "Rest" = "lightblue",
    "Task" = "salmon"
  )) +
  scale_x_continuous(breaks = c(1,2,3,4)) +
  coord_cartesian(ylim = c(3,5)) +
  labs(
    title = "Day 3 - Dominant tremor frequency across repetitions",
    x = "Repetition",
    y = "Dominant frequency (Hz)",
    color = "Condition"
  ) +
  theme_minimal(base_size = 10)
```


Dominant tremor frequency during Day 3 remained relatively stable across repetitions and conditions.

Task frequencies remained within the expected cerebellar tremor frequency range, supporting short-term reproducibility of smartphone-based frequency measurements.



1.3.3 RMS amplitude evolution across repetitions

Objective: To evaluate tremor intensity modulation during Day 3.

RMS amplitude reflects tremor intensity rather than tremor frequency.

```{r}
ggplot(day3_data,
       aes(x = Repetition,
           y = RMS_Amplitude,
           color = Condition,
           group = Condition)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 4) +
  geom_text(
    aes(label = round(RMS_Amplitude, 2)),
    vjust = -1,
    size = 4,
    color = "black"
  ) +
  scale_color_manual(values = c(
    "Rest" = "lightblue",
    "Task" = "salmon"
  )) +
  scale_x_continuous(breaks = c(1,2,3,4)) +
  coord_cartesian(ylim = c(0,1.2)) +
  labs(
    title = "Day 3 - RMS amplitude across repetitions",
    x = "Repetition",
    y = "RMS amplitude",
    color = "Condition"
  ) +
  theme_minimal(base_size = 10)
```


RMS amplitudes remained substantially higher during task conditions than during rest periods, indicating increased tremor intensity during voluntary movement.

This observation is consistent with the action-related characteristics of cerebellar tremor.



1.3.4 Coefficient of variation (CV)

Objective: To evaluate measurement variability across repetitions.

The coefficient of variation (CV) was calculated for:

* dominant tremor frequency,
* RMS amplitude.

A low CV indicates good reproducibility and low variability between repetitions.
```{r}

day3_cv <- day3_data %>%
  group_by(Condition) %>%
  summarise(
    mean_frequency = mean(Dominant_Frequency_Hz),
    sd_frequency = sd(Dominant_Frequency_Hz),
    CV_frequency = (sd_frequency / mean_frequency) * 100,
    mean_RMS = mean(RMS_Amplitude),
    sd_RMS = sd(RMS_Amplitude),
    CV_RMS = (sd_RMS / mean_RMS) * 100,
    .groups = "drop"
  )
day3_cv
```
The coefficient of variation remained relatively low for dominant frequency during task conditions, suggesting acceptable reproducibility across repeated measurements.

Higher RMS variability was observed during rest periods compared with task conditions.



1.3.5 Day 3 synthesis

Overall, Day 3 analysis supports the reproducibility of smartphone-based accelerometry measurements for cerebellar tremor frequency quantification.

Dominant tremor frequency remained within the expected cerebellar tremor range and showed relatively stable values across repetitions.

RMS amplitude remained substantially higher during voluntary movement, supporting the clinical coherence of the measurements obtained using smartphone accelerometry and FFT analysis.

















































































