---
title: "Statistical analysis of cerebellar tremor frequency"
author: 'Sofia ZEMMAHI '
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output:
  html_document:
    df_print: paged
    code_folding: hide
    toc: yes
    number_sections: yes
  word_document:
    toc: yes
  pdf_document:
    toc: yes
---

**# Objective**

To evaluate whether the smartphone accelerometer application *PhyPhox* can reproducibly measure cerebellar tremor characteristics.

More specifically, this study aimed to determine whether dominant tremor frequencies remained consistent:

* between repetitions,
* between recording days,
* and between rest and task conditions.

The main variable of interest was dominant tremor frequency, as cerebellar tremor is classically characterized by low-frequency oscillations ranging from 3 to 5 Hz.

Fast Fourier Transform (FFT) analysis was used to extract dominant tremor frequency from accelerometer recordings.

The primary outcome measure was therefore **Dominant_Frequency_Hz**.

A secondary variable of interest was **RMS amplitude** (Root Mean Square amplitude), which reflects tremor intensity rather than tremor frequency.

RMS amplitude was used to:

* quantify tremor intensity,
* evaluate whether tremor increased during voluntary movement,
* and reinforce the clinical consistency of the measurements.
The main question of this analysis was therefore:

**Can the smartphone application *PhyPhox* provide reproducible tremor measurements consistent with cerebellar tremor characteristics?**

This question was explored through three sub-parts:

1. Do dominant frequencies remain stable across repetitions?  
→ Intra-day reproducibility

2. Do dominant frequencies remain stable across recording days?  
→ Inter-day reproducibility

3. Does tremor intensity increase during voluntary movement?

**# Part 1 : Intra-day reproducibility**

## Do dominant tremor frequencies remain stable between repetitions?

The objective of this first part was to evaluate short-term reproducibility of dominant tremor frequency measurements obtained using smartphone accelerometry and FFT analysis.

For each recording day, dominant tremor frequency was analyzed across repeated measurements performed during rest and task conditions.

A stable frequency between repetitions would support the reproducibility of the measurement method.







## 1.1.4 Coefficient of variation (CV)

Objective: To evaluate measurement variability across repetitions.

The coefficient of variation (CV) was calculated for:

- dominant tremor frequency,

- RMS amplitude.

A low CV indicates good reproducibility and low variability between repetitions.

During task conditions, dominant frequency showed a low coefficient of variation (CV = 1.73%), suggesting good reproducibility across repetitions.

Conversely, RMS amplitude during rest showed higher variability (CV = 82.27%), reflecting fluctuations in tremor intensity during resting periods.

## 1.1.5 Day 1 synthesis

Overall, Day 1 analysis suggests that smartphone accelerometry combined with FFT analysis provides coherent dominant tremor frequency measurements consistent with cerebellar tremor characteristics.

Dominant frequency remained within the classical cerebellar tremor range (3–5 Hz) and showed relatively good stability across task repetitions.

RMS analysis demonstrated that tremor intensity increased substantially during voluntary movement compared with rest conditions.

These findings support the feasibility of smartphone-based tremor frequency quantification during repeated measurements.

```{r}
library(readr)
all_results <- read_csv("final_data/all_results.csv")
View(all_results)
```
# 1.1 Day 1 analysis

## 1.1.1 Descriptive statistics

Objective: To describe dominant tremor frequencies obtained during Day 1.

We calculated:

- mean frequency,
- standard deviation (SD),
- median,
- minimum frequency,
- maximum frequency.

These descriptive statistics help verify whether the obtained frequencies remained within the classical cerebellar tremor range reported in the literature (3–5 Hz).
Dominant tremor frequency during Day 1 remained within the classical cerebellar tremor range for both rest and task conditions.

```{r day1_data, message=FALSE, warning=FALSE}
# Select Day 1 data
day1_data <- all_results[all_results$Day == "D1", ]

```

Descriptive statistics on day 1 
```{r day1_stats, message=FALSE, warning=FALSE}

# Mean frequency
mean_frequency <- tapply(
  day1_data$Dominant_Frequency_Hz,
  day1_data$Condition,
  mean
)

# Standard deviation
sd_frequency <- tapply(
  day1_data$Dominant_Frequency_Hz,
  day1_data$Condition,
  sd
)

# Median
median_frequency <- tapply(
  day1_data$Dominant_Frequency_Hz,
  day1_data$Condition,
  median
)

# Minimum
min_frequency <- tapply(
  day1_data$Dominant_Frequency_Hz,
  day1_data$Condition,
  min
)

# Maximum
max_frequency <- tapply(
  day1_data$Dominant_Frequency_Hz,
  day1_data$Condition,
  max
)

# Final table
day1_stats <- data.frame(
  Condition = names(mean_frequency),
  mean_frequency,
  sd_frequency,
  median_frequency,
  min_frequency,
  max_frequency
)

day1_stats

```


Dominant tremor frequency during Day 1 remained within the classical cerebellar tremor range (3–5 Hz) for both rest and task conditions.



## 1.1.2 Dominant frequency evolution across repetitions

Objective: To evaluate frequency stability across repetitions during Day 1.

A line plot was used to visualize frequency evolution over repeated measurements.
```{r}
library(ggplot2)
library(dplyr)
library(readr)
ggplot(day1_data,
       aes(x = Repetition,
           y = Dominant_Frequency_Hz,
           color = Condition,
           group = Condition)) +

  geom_line(linewidth = 1.2) +

  geom_point(size = 4) +

  geom_text(
    aes(label = round(Dominant_Frequency_Hz, 2)),
    vjust = -1,
    size = 4,
    color = "black"
  ) +

  scale_color_manual(values = c(
    "Rest" = "lightblue",
    "Task" = "salmon"
  )) +

  scale_x_continuous(breaks = c(1,2,3,4)) +

  coord_cartesian(ylim = c(3,5)) +

  labs(
    title = "Day 1 - Dominant tremor frequency across repetitions",
    x = "Repetition",
    y = "Dominant frequency (Hz)",
    color = "Condition"
  ) +

  theme_minimal(base_size = 10)
```


During task conditions, dominant tremor frequency remained relatively stable across repetitions (3.72 Hz and 3.63 Hz), suggesting acceptable short-term reproducibility.

Greater variability was observed during rest conditions, with frequencies increasing from 3.27 Hz to 4.69 Hz and 4.85 Hz across repetitions.

Despite these fluctuations, all frequency values remained within the classical cerebellar tremor range reported in the literature (3–5 Hz).



1.1.3 RMS amplitude evolution across repetitions

Objective: To evaluate tremor intensity modulation during Day 1.

RMS amplitude reflects tremor intensity rather than tremor frequency.

```{r}
ggplot(day1_data,
       aes(x = Repetition,
           y = RMS_Amplitude,
           color = Condition,
           group = Condition)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 4) +
  geom_text(
    aes(label = round(RMS_Amplitude, 2)),
    vjust = -1,
    size = 4,
    color = "black"
  ) +
  scale_color_manual(values = c(
    "Rest" = "lightblue",
    "Task" = "salmon"
  )) +
  scale_x_continuous(breaks = c(1,2,3,4)) +
  coord_cartesian(ylim = c(0,1)) +
  labs(
    title = "Day 1 - RMS amplitude across repetitions",
    x = "Repetition",
    y = "RMS amplitude",
    color = "Condition"
  ) +
  theme_minimal(base_size = 10)

```
RMS amplitude values were substantially higher during task conditions compared with rest conditions.

During task repetitions, RMS amplitude reached approximately 0.69 and 0.72, whereas rest values remained markedly lower overall.

These findings suggest increased tremor intensity during voluntary movement, which is clinically consistent with cerebellar tremor behavior.


1.1.4 Coefficient of variation (CV)

Objective: To evaluate measurement variability across repetitions.

The coefficient of variation (CV) was calculated for:

* dominant tremor frequency,
* RMS amplitude.

A low CV indicates good reproducibility and low variability between repetitions.

```{r}
day1_cv <- day1_data %>%
  group_by(Condition) %>%
  summarise(
    mean_frequency = mean(Dominant_Frequency_Hz),
    sd_frequency = sd(Dominant_Frequency_Hz),
    CV_frequency = (sd_frequency / mean_frequency) * 100,
    mean_RMS = mean(RMS_Amplitude),
    sd_RMS = sd(RMS_Amplitude),
    CV_RMS = (sd_RMS / mean_RMS) * 100,
    .groups = "drop"
  )
day1_cv

```
During task conditions, dominant frequency showed a low coefficient of variation (CV = 1.73%), suggesting good reproducibility across repetitions.

Conversely, RMS amplitude during rest showed higher variability (CV = 82.27%), reflecting fluctuations in tremor intensity during resting periods.


1.1.5 Day 1 synthesis

Overall, Day 1 analysis suggests that smartphone accelerometry combined with FFT analysis provides coherent dominant tremor frequency measurements consistent with cerebellar tremor characteristics.

Dominant frequency remained within the classical cerebellar tremor range (3–5 Hz) and showed relatively good stability across task repetitions.

RMS analysis demonstrated that tremor intensity increased substantially during voluntary movement compared with rest conditions.

These findings support the feasibility of smartphone-based tremor frequency quantification during repeated measurements.





# 1.2 Day 2 analysis

## 1.2.1 Descriptive statistics

Objective: To describe dominant tremor frequencies obtained during Day 2.

We calculated:
- mean frequency,
- standard deviation (SD),
- median,
- minimum frequency,
- maximum frequency.

These descriptive statistics help verify whether the obtained frequencies remain within the classical cerebellar tremor range reported in the literature (3–5 Hz).

```{r}
# Select Day 2 data

day2_data <- all_results %>%
  filter(Day == "D2")

# Descriptive statistics for Day 2

day2_stats <- day2_data %>%
  group_by(Condition) %>%
  summarise(
    mean_frequency = mean(Dominant_Frequency_Hz),
    sd_frequency = sd(Dominant_Frequency_Hz),
    median_frequency = median(Dominant_Frequency_Hz),
    min_frequency = min(Dominant_Frequency_Hz),
    max_frequency = max(Dominant_Frequency_Hz),
    .groups = "drop"
  )

day2_stats
```
Dominant tremor frequency during Day 2 remained within the classical cerebellar tremor range (3–5 Hz) for both rest and task conditions.

1.2.2 Dominant frequency evolution across repetitions

Objective: To evaluate frequency stability across repetitions during Day 2.

A line plot was used to visualize frequency evolution over repeated measurements.
```{r}
ggplot(day2_data,

       aes(x = Repetition,

           y = Dominant_Frequency_Hz,

           color = Condition,

           group = Condition)) +

  geom_line(linewidth = 1.2) +

  geom_point(size = 4) +

  geom_text(

    aes(label = round(Dominant_Frequency_Hz, 2)),

    vjust = -1,

    size = 4,

    color = "black"

  ) +

  scale_color_manual(values = c(

    "Rest" = "lightblue",

    "Task" = "salmon"

  )) +

  scale_x_continuous(breaks = c(1,2,3,4)) +

  coord_cartesian(ylim = c(3,5)) +

  labs(

    title = "Day 2 - Dominant tremor frequency across repetitions",

    x = "Repetition",

    y = "Dominant frequency (Hz)",

    color = "Condition"

  ) +

  theme_minimal(base_size = 10)
```
Dominant tremor frequency during Day 2 remained globally stable across repetitions and conditions.

Task frequencies remained within the expected cerebellar tremor frequency range, supporting short-term reproducibility of frequency measurements.



1.2.3 RMS amplitude evolution across repetitions

Objective: To evaluate tremor intensity modulation during Day 2.

RMS amplitude reflects tremor intensity rather than tremor frequency.
```{r}
ggplot(day2_data,

       aes(x = Repetition,

           y = RMS_Amplitude,

           color = Condition,

           group = Condition)) +

  geom_line(linewidth = 1.2) +

  geom_point(size = 4) +

  geom_text(

    aes(label = round(RMS_Amplitude, 2)),

    vjust = -1,

    size = 4,

    color = "black"

  ) +

  scale_color_manual(values = c(

    "Rest" = "lightblue",

    "Task" = "salmon"

  )) +

  scale_x_continuous(breaks = c(1,2,3,4)) +

  coord_cartesian(ylim = c(0,1.2)) +

  labs(

    title = "Day 2 - RMS amplitude across repetitions",

    x = "Repetition",

    y = "RMS amplitude",

    color = "Condition"

  ) +

  theme_minimal(base_size = 10)
```
RMS amplitudes remained substantially higher during task conditions than during rest periods, indicating increased tremor intensity during voluntary movement.
This observation is consistent with the action-related characteristics of cerebellar tremor.


1.2.4 Coefficient of variation (CV)

Objective: To evaluate measurement variability across repetitions.

The coefficient of variation (CV) was calculated for:

* dominant tremor frequency,
* RMS amplitude.

A low CV indicates good reproducibility and low variability between repetitions.
```{r}
day2_cv <- day2_data %>%

  group_by(Condition) %>%

  summarise(

    mean_frequency = mean(Dominant_Frequency_Hz),

    sd_frequency = sd(Dominant_Frequency_Hz),

    CV_frequency = (sd_frequency / mean_frequency) * 100,

    mean_RMS = mean(RMS_Amplitude),

    sd_RMS = sd(RMS_Amplitude),

    CV_RMS = (sd_RMS / mean_RMS) * 100,

    .groups = "drop"

  )

day2_cv
```

The coefficient of variation remained relatively low for dominant frequency during task conditions, suggesting good reproducibility across repeated measurements.

Higher RMS variability was observed during rest periods compared with task conditions.


1.2.5 Day 2 synthesis

Overall, Day 2 analysis supports the reproducibility of smartphone-based accelerometry measurements for cerebellar tremor frequency quantification.

Dominant tremor frequency remained within the expected cerebellar tremor range and showed relatively stable values across repetitions.

RMS amplitude remained substantially higher during voluntary movement, supporting the clinical coherence of the measurements obtained using smartphone accelerometry and FFT analysis.



# 1.3 Day 3 analysis

## 1.3.1 Descriptive statistics

Objective: To describe dominant tremor frequencies obtained during Day 3.

We calculated:
- mean frequency,
- standard deviation (SD),
- median,
- minimum frequency,
- maximum frequency.

These descriptive statistics help verify whether the obtained frequencies remain within the classical cerebellar tremor range reported in the literature (3–5 Hz).

```{r}
# Select Day 3 data

day3_data <- all_results %>%
  filter(Day == "D3")

# Descriptive statistics for Day 3
day3_stats <- day3_data %>%
  group_by(Condition) %>%
  summarise(
    mean_frequency = mean(Dominant_Frequency_Hz),
    sd_frequency = sd(Dominant_Frequency_Hz),
    median_frequency = median(Dominant_Frequency_Hz),
    min_frequency = min(Dominant_Frequency_Hz),
    max_frequency = max(Dominant_Frequency_Hz),
    .groups = "drop"
  )

day3_stats
```

Dominant tremor frequency during Day 3 remained within the classical cerebellar tremor range (3–5 Hz) for both rest and task conditions.

1.3.2 Dominant frequency evolution across repetitions

Objective: To evaluate frequency stability across repetitions during Day 3.

A line plot was used to visualize frequency evolution over repeated measurements.

```{r}
ggplot(day3_data,
       aes(x = Repetition,
           y = Dominant_Frequency_Hz,
           color = Condition,
           group = Condition)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 4) +
  geom_text(
    aes(label = round(Dominant_Frequency_Hz, 2)),
    vjust = -1,
    size = 4,
    color = "black"
  ) +
  scale_color_manual(values = c(
    "Rest" = "lightblue",
    "Task" = "salmon"
  )) +
  scale_x_continuous(breaks = c(1,2,3,4)) +
  coord_cartesian(ylim = c(3,5)) +
  labs(
    title = "Day 3 - Dominant tremor frequency across repetitions",
    x = "Repetition",
    y = "Dominant frequency (Hz)",
    color = "Condition"
  ) +
  theme_minimal(base_size = 10)
```


Dominant tremor frequency during Day 3 remained relatively stable across repetitions and conditions.

Task frequencies remained within the expected cerebellar tremor frequency range, supporting short-term reproducibility of smartphone-based frequency measurements.



1.3.3 RMS amplitude evolution across repetitions

Objective: To evaluate tremor intensity modulation during Day 3.

RMS amplitude reflects tremor intensity rather than tremor frequency.

```{r}
ggplot(day3_data,
       aes(x = Repetition,
           y = RMS_Amplitude,
           color = Condition,
           group = Condition)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 4) +
  geom_text(
    aes(label = round(RMS_Amplitude, 2)),
    vjust = -1,
    size = 4,
    color = "black"
  ) +
  scale_color_manual(values = c(
    "Rest" = "lightblue",
    "Task" = "salmon"
  )) +
  scale_x_continuous(breaks = c(1,2,3,4)) +
  coord_cartesian(ylim = c(0,1.2)) +
  labs(
    title = "Day 3 - RMS amplitude across repetitions",
    x = "Repetition",
    y = "RMS amplitude",
    color = "Condition"
  ) +
  theme_minimal(base_size = 10)
```


RMS amplitudes remained substantially higher during task conditions than during rest periods, indicating increased tremor intensity during voluntary movement.

This observation is consistent with the action-related characteristics of cerebellar tremor.



1.3.4 Coefficient of variation (CV)

Objective: To evaluate measurement variability across repetitions.

The coefficient of variation (CV) was calculated for:

* dominant tremor frequency,
* RMS amplitude.

A low CV indicates good reproducibility and low variability between repetitions.
```{r}

day3_cv <- day3_data %>%
  group_by(Condition) %>%
  summarise(
    mean_frequency = mean(Dominant_Frequency_Hz),
    sd_frequency = sd(Dominant_Frequency_Hz),
    CV_frequency = (sd_frequency / mean_frequency) * 100,
    mean_RMS = mean(RMS_Amplitude),
    sd_RMS = sd(RMS_Amplitude),
    CV_RMS = (sd_RMS / mean_RMS) * 100,
    .groups = "drop"
  )
day3_cv
```
The coefficient of variation remained relatively low for dominant frequency during task conditions, suggesting acceptable reproducibility across repeated measurements.

Higher RMS variability was observed during rest periods compared with task conditions.


1.3.5 Day 3 synthesis

Overall, Day 3 analysis supports the reproducibility of smartphone-based accelerometry measurements for cerebellar tremor frequency quantification.

Dominant tremor frequency remained within the expected cerebellar tremor range and showed relatively stable values across repetitions.

RMS amplitude remained substantially higher during voluntary movement, supporting the clinical coherence of the measurements obtained using smartphone accelerometry and FFT analysis.



**# Part 2 : Inter-day reproducibility**

## Do dominant tremor frequencies remain consistent between recording days?

The objective of this second part is to evaluate inter-day reproducibility of dominant tremor frequency measurements obtained using smartphone accelerometry and FFT analysis.
To answer this question, dominant tremor frequency values obtained during Day 1, Day 2 and Day 3 were compared.
Stable frequency values across recording days would support the reproducibility of smartphone-based tremor quantification.


# 2.1 Mean dominant frequency across recording days
Objective: To compare mean dominant tremor frequency between Day 1, Day 2 and Day 3.

Mean frequency values were calculated separately for:
- rest conditions,
- task conditions.
```{r}
daily_frequency <- all_results %>%
  group_by(Day, Condition) %>%
  summarise(
    mean_frequency = mean(Dominant_Frequency_Hz),
    sd_frequency = sd(Dominant_Frequency_Hz),
    .groups = "drop"
  )
daily_frequency
```
The obtained frequencies remained within the classical cerebellar tremor range (3–5 Hz) across all recording days.


2.2 Evolution of mean dominant frequency across days

Objective: To visualize frequency stability between recording days.

```{r}
ggplot(daily_frequency,
       aes(x = Day,
           y = mean_frequency,
           color = Condition,
           group = Condition)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 4) +
  geom_errorbar(
    aes(
      ymin = mean_frequency - sd_frequency,
      ymax = mean_frequency + sd_frequency
    ),
    width = 0.1
  ) +
  geom_text(
    aes(label = round(mean_frequency, 2)),
    vjust = -1,
    size = 4,
    color = "black"
  ) +
  scale_color_manual(values = c(
    "Rest" = "lightblue",
    "Task" = "salmon"
  )) +
  coord_cartesian(ylim = c(3,5)) +
  labs(
    title = "Mean dominant tremor frequency across days",
    x = "Recording day",
    y = "Mean dominant frequency (Hz)",
    color = "Condition"
  ) +
  theme_minimal(base_size = 10)

```
Mean dominant tremor frequency remained relatively stable across recording days for both rest and task conditions.

Only limited frequency variations were observed between days, supporting acceptable inter-day reproducibility.


2.3 Inter-day coefficient of variation

Objective: To evaluate variability of dominant frequency measurements between recording days.

The coefficient of variation (CV) was calculated across recording days.

A low CV indicates better inter-day reproducibility.

```{r}
interday_cv <- daily_frequency %>%
  group_by(Condition) %>%
  summarise(
    overall_mean_frequency = mean(mean_frequency),
    overall_sd_frequency = sd(mean_frequency),
    interday_CV =
      (overall_sd_frequency / overall_mean_frequency) * 100,
    .groups = "drop"
  )
interday_cv
```


Low inter-day CV values suggest relatively stable dominant frequency measurements across recording sessions.

2.4 Exploratory statistical comparison between days

Objective: To explore whether dominant tremor frequency differed significantly between recording days.

Because this study included only one patient and a limited number of repetitions, a non-parametric exploratory approach was preferred.
```{r}

kruskal.test(
  Dominant_Frequency_Hz ~ Day,
  data = all_results
)

```

The Kruskal-Wallis test was used as an exploratory analysis to compare dominant tremor frequency between recording days.

Given the exploratory single-patient design, these results should be interpreted cautiously.


2.5 Inter-day reproducibility synthesis

Overall, dominant tremor frequency remained within the expected cerebellar tremor range across Day 1, Day 2 and Day 3.

Mean frequencies showed relatively small variations between recording days, and inter-day coefficient of variation values remained acceptable.

These findings suggest that smartphone accelerometry combined with FFT analysis may provide reproducible dominant tremor frequency measurements across repeated recording sessions.

However, these observations remain exploratory and should be confirmed in larger patient cohorts.




**# Part 3 : Effect of action on tremor intensity**

## Does tremor intensity increase during voluntary movement?

The objective of this third part is to evaluate whether tremor intensity increases during voluntary movement compared with rest conditions.

To answer this question, RMS amplitude values obtained during rest and task conditions were compared.
Unlike dominant frequency, which reflects the rhythmic component of tremor, RMS amplitude reflects the intensity of the movement signal.
An increase in RMS amplitude during task conditions would therefore support the action-related characteristics of cerebellar tremor.

# 3.1 Mean RMS amplitude according to condition

Objective: To compare mean RMS amplitude between rest and task conditions.
```{r}
rms_condition_stats <- all_results %>%
  group_by(Condition) %>%
  summarise(
    mean_RMS = mean(RMS_Amplitude),
    sd_RMS = sd(RMS_Amplitude),
    median_RMS = median(RMS_Amplitude),
    min_RMS = min(RMS_Amplitude),
    max_RMS = max(RMS_Amplitude),
    .groups = "drop"
  )
rms_condition_stats
```

Mean RMS amplitude was substantially higher during task conditions compared with rest periods.



3.2 RMS amplitude evolution during rest and task conditions

Objective: To visualize differences in tremor intensity between rest and task conditions.

```{r}
ggplot(all_results,
       aes(x = Condition,
           y = RMS_Amplitude,
           fill = Condition)) +
  geom_boxplot(alpha = 0.7) +
  geom_jitter(width = 0.1,
              size = 3,
              alpha = 0.8) +
  scale_fill_manual(values = c(
    "Rest" = "lightblue",
    "Task" = "salmon"
  )) +
  labs(
    title = "RMS amplitude according to condition",
    x = "Condition",
    y = "RMS amplitude"
  ) +
  theme_minimal(base_size = 10)
```

RMS amplitude values were globally higher during task conditions, suggesting increased tremor intensity during voluntary movement.


3.3 Exploratory statistical comparison between rest and task conditions

Objective: To explore whether RMS amplitude differed between rest and task conditions.

Because this study included a single patient and a limited number of repetitions, a non-parametric exploratory approach was preferred.

```{r}
wilcox.test(
  RMS_Amplitude ~ Condition,
  data = all_results
)

```
The Wilcoxon test showed a significant difference in RMS amplitude between rest and task conditions (p < 0.001), with substantially higher RMS values observed during voluntary movement.

These findings are consistent with the classical clinical characteristics of cerebellar tremor, which typically increases during action and goal-directed movement.

However, these results should be interpreted cautiously. The present work was designed as an exploratory feasibility and proof-of-concept study based on a single patient and a limited number of repetitions.

Therefore, the objective was not to demonstrate a definitive physiological effect, but rather to evaluate whether smartphone accelerometry combined with FFT analysis could detect coherent and clinically meaningful tremor characteristics.

Further studies including larger patient cohorts and standardized acquisition protocols would be necessary to confirm the reproducibility and clinical relevance of these preliminary observations.


3.4 RMS amplitude according to recording day

Objective: To visualize whether the increase in tremor intensity during movement remained consistent across recording days.

```{r}
daily_rms <- all_results %>%
  group_by(Day, Condition) %>%
  summarise(
    mean_RMS = mean(RMS_Amplitude),
    sd_RMS = sd(RMS_Amplitude),
    .groups = "drop"
  )
daily_rms
ggplot(daily_rms,
       aes(x = Day,
           y = mean_RMS,
           color = Condition,
           group = Condition)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 4) +
  geom_errorbar(
    aes(
      ymin = mean_RMS - sd_RMS,
      ymax = mean_RMS + sd_RMS
    ),
    width = 0.1
  ) +
  geom_text(
    aes(label = round(mean_RMS, 2)),
    vjust = -1,
    size = 4,
    color = "black"
  ) +
  scale_color_manual(values = c(
    "Rest" = "lightblue",
    "Task" = "salmon"
  )) +
  coord_cartesian(ylim = c(0,1.2)) +
  labs(
    title = "Mean RMS amplitude across days",
    x = "Recording day",
    y = "Mean RMS amplitude",
    color = "Condition"
  ) +
  theme_minimal(base_size = 10)
```


Across all recording days, RMS amplitude remained consistently higher during task conditions than during rest periods.

This observation supports the action-related nature of cerebellar tremor.


3.5 Part 3 synthesis

Overall, RMS amplitude values were substantially higher during voluntary movement than during rest conditions.

These findings support the clinical characteristics of cerebellar tremor, which typically increases during action and goal-directed movement.

The consistency of RMS elevation across recording days suggests that smartphone accelerometry may provide clinically relevant information not only about tremor frequency, but also about tremor intensity during movement.


**# Global synthesis of the results**

Overall, the analyses performed in this study suggest that smartphone accelerometry combined with FFT analysis can provide coherent measurements of cerebellar tremor characteristics in a proof-of-concept framework.

Across the three recording days, dominant tremor frequencies consistently remained within the classical cerebellar tremor range described in the literature (approximately 3–5 Hz). Frequency values showed relatively limited variability between repetitions and between recording days, supporting acceptable intra-day and inter-day reproducibility of the measurements.

During task conditions, dominant tremor frequency generally remained stable across repetitions, whereas greater variability was observed during some rest periods. These fluctuations may reflect residual oscillatory activity, postural adjustments, involuntary corrective movements, or movement artefacts occurring during resting phases.

RMS amplitude analysis showed a markedly different behavior. RMS values were consistently and substantially higher during voluntary movement compared with rest conditions across all recording days. This observation is clinically coherent with cerebellar tremor physiology, as cerebellar tremor classically increases during action and goal-directed movement.

The exploratory Wilcoxon analysis also suggested a difference in RMS amplitude between rest and task conditions. However, these statistical findings should be interpreted cautiously due to the exploratory proof-of-concept design, the single-patient dataset, and the limited number of repetitions.

Taken together, these results support the feasibility of using smartphone accelerometry to characterize cerebellar tremor frequency and tremor intensity in a simple and accessible manner.

Although the present work does not aim to establish definitive clinical conclusions, it provides preliminary evidence supporting the potential utility of smartphone-based tremor analysis tools in future larger-scale studies.

