---
title: "Statistical analysis of cerebellar tremor frequency"
author: 'Sofia ZEMMAHI '
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output:
  html_document:
    df_print: paged
    code_folding: hide
    toc: yes
    number_sections: yes
  pdf_document:
    toc: yes
---

# Objective

The objective of this R analysis is to explore the reproducibility of smartphone-based accelerometry for quantifying cerebellar tremor frequency across repeated grasping tasks and across three consecutive days.

The analysis is based on the dominant tremor frequency and RMS amplitude extracted from Python using FFT.



# Import dataframe from the final data folder

```{r}
library(tidyverse)
all_results <- read.csv("final_data/all_results.csv")
View(all_results)
```


# Overview of the dataset

The final dataframe contains the dominant tremor frequency and RMS amplitude extracted from each rest and grasping phase across the three experimental days.
```{r}

head(all_results)

summary(all_results)

str(all_results)

```

#  Descriptive statistics

Descriptive statistics were computed to summarize dominant tremor frequency and RMS amplitude across days and repetitions.

Mean values, standard deviations, medians, and ranges were calculated in order to obtain a first overview of tremor variability and measurement stability.

```{r}
library(tidyverse)

descriptive_stats <- all_results %>%
  group_by(Day, Condition) %>%
  summarise(
    mean_frequency = mean(Dominant_Frequency_Hz),
    sd_frequency = sd(Dominant_Frequency_Hz),
    median_frequency = median(Dominant_Frequency_Hz),
    min_frequency = min(Dominant_Frequency_Hz),
    max_frequency = max(Dominant_Frequency_Hz),
    mean_rms = mean(RMS_Amplitude),
    sd_rms = sd(RMS_Amplitude),
    .groups = "drop"
  )

descriptive_stats
```

