---
title: "Statistical analysis of cerebellar tremor frequency"
author: 'Sofia ZEMMAHI '
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
output:
  html_document:
    df_print: paged
    code_folding: hide
    toc: yes
    number_sections: yes
  pdf_document:
    toc: yes
---

# Objective

To verify if a smartphone accelerometre *PhyPhox* can measure cerebellar tremor reproducibly.

And More ore specifically:
Are the obtained frequencies consistent:

* between repetitions,
* between days,
* and between conditions?

The main variable of this study is the dominant frequency because cerebellar tremor is defined by a low frequency (<5 Hz),
and the use of FFT to extract that frequency.

Here the main outcome is Dominant_Frequency_Hz.

A secondary essential variable is RMS wich is the Root Mean square amplitude. 
It measures the intensity of the tremor, not the frequent.
It is used to: quantify the intensity, verify that the tremor increases during the action and strengthen clinical consistency.

So the main question of this analysis is : 
**Does the smartphone *PhyPhox* accelerometre measure a cohent frequency with cerebellar tremor?**

We can answer this principal question in three under-parts : 
1) Do the frequencies remain stable between repetitions? -> Intra day reproducibility.
2) Do the frequencies remain consistent between days? -> Inter day reproducibility.
3) Does tremor increase during action?


# Part 1 : Dominant Frequency 

1. Descriptive Statistics

Objective: To describe the obtained frequencies.

We calculate : 
* mean frequency
* SD
* min/max
* median

Why ? 
To verify if frequencies are in cerebellar area ans and if they are close between repetitions.


# Import dataframe from the final data folder

```{r}

# Import dataframe from the final data folder
library(tidyverse)

all_results <- read.csv("final_data/all_results.csv")

View(all_results)

```


```{r}
frequency_stats <- all_results %>%
  group_by(Day, Condition) %>%
  summarise(
    mean_frequency = mean(Dominant_Frequency_Hz),
    sd_frequency = sd(Dominant_Frequency_Hz),
    median_frequency = median(Dominant_Frequency_Hz),
    min_frequency = min(Dominant_Frequency_Hz),
    max_frequency = max(Dominant_Frequency_Hz),
    .groups = "drop"
  )

frequency_stats
```
Dominant tremor frequency remained within the classical cerebellar tremor range (3–5 Hz) across repetitions and days.


# Part 2 : Evolution over time 

We want to see stability, reproducibility and trend.

The most appropriate is a LINE plot.

1. Plot for day 1 : Dominant frequency 
```{r}
# Select Day 1 data
day1_data <- all_results %>%
  filter(Day == "D1")

ggplot(day1_data,
       aes(x = Repetition,
           y = Dominant_Frequency_Hz,
           color = Condition,
           group = Condition)) +
  
  geom_line(linewidth = 1.2) +
  geom_point(size = 4) +
  
  geom_text(
    aes(label = round(Dominant_Frequency_Hz, 2)),
    vjust = -1,
    size = 4,
    color = "black"
  ) +
  
  scale_color_manual(values = c(
    "Rest" = "lightblue",
    "Task" = "salmon"
  )) +
  
  scale_x_continuous(breaks = c(1, 2, 3, 4)) +
  coord_cartesian(ylim = c(3, 5)) +
  
  labs(
    title = "Day 1 - Dominant tremor frequency across repetitions",
    x = "Repetition",
    y = "Dominant frequency (Hz)",
    color = "Condition"
  ) +
  
  theme_minimal(base_size = 10)
```

2.Plot 2 dor day 1 : RMS with rest + Task 
```{r}
day1_data <- all_results %>%

  filter(Day == "D1")

ggplot(day1_data,
       aes(x = Repetition,
           y = RMS_Amplitude,
           color = Condition,
           group = Condition)) +
  
  geom_line(linewidth = 1.2) +
  geom_point(size = 4) +
  geom_text(
    aes(label = round(RMS_Amplitude, 2)),
    vjust = -1,
    size = 4,
    color = "black"
  ) +

  scale_color_manual(values = c(
    "Rest" = "lightblue",
    "Task" = "salmon"
  )) +

  scale_x_continuous(
    breaks = c(1, 2, 3, 4)
  ) +
  
  # Zoomed RMS scale
  coord_cartesian(
    ylim = c(0, 1)
  ) +
  labs(
    title = "Day 1 - RMS amplitude across repetitions",
    x = "Repetition",
    y = "RMS amplitude",
    color = "Condition"
  ) +
  
  theme_minimal(base_size = 10)
```











